\bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in Drosophila

### 2018-02-07 latex/paper.tex
updated

In the male Drosophila germline, nascent TZs first assemble on basal bodies
during early pre-meiotic G2 phase (spermatocytes), anchor directly to the plasma membrane.
This is followed by establishment of cilium-membrane connections and a
ciliary membrane that persists during meiosis \citep{riparbelli2012assembly}
through a poorly understood mechanism
involving microtubule rearrangements \citep{gottardo2013cilium}.
Ciliogenesis begins with the assembly of TZ proteins at the tips of
the nascent basal body.
This is followed by the anchoring of the TZ to the membrane and
establishment of a compartmentalized space bounded by the ciliary membrane
and the TZ.
TZ maturation has been described in
\textit{Paramecium} \citep{aubusson2015transition}
\textit{Caenorrhabditis elegans} \citep{serwas2017centrioles} and
\textit{D. melanogaster} \citep{gottardo2013cilium}
and involves TZ elongation.
Possibly, this is similar to the initial steps of TZ assembly and reorganization
in well-studied vertebrate cilia.

Here, we show that the lipid phosphatidylinositol 4,5-bisphosphate (\PIP{})
is essential for proper TZ maturation in developing \textit{Drosophila melanogaster}
male germ cells.
Reduction of cellular \PIP{} by expression of the bacterial PIP phosphatase SigD
or by inactivation of the PIP 5-kinase Skittles (Sktl) that localizes to cilia
induces longer than normal TZs.
These hyperelongated TZs show hallmarks of aberrant function similar to those
found in previously studied TZ mutants, including loss of cilium-membrane associations.
We also find that the \textit{onion rings} allele of
ciliopathy-associated exocyst protein Exo84/EXOC8 decouples TZ hyperelongation
from loss of cilium-membrane contacts.
Our results suggest that the plasma membrane lipid \PIP{} regulates TZ length,
maturation and function, acting to support ciliogenesis.

@@

Since exocyst regulates membrane trafficking and cilium formation,
we examined whether cilium-associated membrane dynamics were affected in \sigd{}
or \textit{onr} mutants similar to \textit{dila; cby} mutants
\citep{vieillard2016transition}.
We found that while TZs in \sigd{} and \textit{onr} cells anchored properly
to the plasma membrane, they were unable to either establish or maintain
membrane connections, and were rendered cytoplasmic (Figure 5) similar to
\textit{dila; cby} mutants.
Although we were not able to determine the localization of the Exo84 due to the
lack of available antibodies,
we found that Sec6 localized to the plasma membrane as expected.
Our results suggest that the exocyst may act downstream of \PIP{} to
enhance cilium-membrane associations, and that TZ hyperelongation and loss of
membrane-cilium association are genetically separable phenotypes.

@@

In this study, we show that the lipid \PIP{} is essential for proper
transition zone maturation in the male Drosophila germline.
Reduction of cellular \PIP{} by expression of the bacterial PIP
phosphatase SigD or inactivation of the type I PIP kinase Skittles
causes developing TZs to elongate excessively.

We had previously shown that reduction of \PIP{} by SigD expression
affected flagellar biogenesis and axoneme structure
\citep{wei2008depletion}.
We also showed aberrant ``comet-shaped'' localization of Unc that
failed to assemble TZ-associated puncta in some cases.
Here, we show that this occurs prior to meiosis,
is not an artefact due to cytokinesis defects
and do not represent an inability of the axoneme to grow.
Rather,

@@

\section{Methods}
\subsection{Transgenic flies}
Drosophila stocks were cultured on cornmeal molasses agar medium at 25$^{\circ}$C
and 50\% humidity.
Stocks expressing $\beta$\textit{2tub}::SigD and $\beta$\textit{2tub}::YFP-Sktl
were described previously \citep{wei2008depletion, fabian2010phosphatidylinositol}.
$\beta$\textit{2tub}::mCherry-Sktl\textsuperscript{kinase-dead}
was generated by Gordon Polevoy.
Ana1-tdTomato (chromosome 2) and Cep290-GFP (chromosome 3) were provided
by T. Avidor-Reiss \citep{basiri2014migrating}.
Unc-GFP was originally provided by M. Kernan \citep{baker2004mechanosensory}.
Stocks expressing GFP-tagged Chibby, B9d1, B9d2, Cc2d2a and Mks1 were generously
provided by B. Durand \citep{enjolras2012drosophila, vieillard2016transition}.
The \textit{onr} mutant was described previously \citep{giansanti2015exocyst}.
Stocks for generating \textit{sktl}\textsuperscript{2.3} clones were originally provided by
A. Guichet \citep{gervais2008pip5k}.
\textit{w}\textsuperscript{1118} was used as the wild-type control.

@@

\subsection{Fluorescence microscopy}
For live imaging, testes were dissected in phosphate bufferd saline (PBS),
transferred to a polylysine-coated glass slide in a drop of PBS,
ruptured using a syringe needle and
squashed under a glass coverslip using Kimwipes.
The edges of the coverslip were sealed with nail polish
and the specimen was visualized immediately using a fluorescence microscope.

>>>>

In the Drosophila male germline, nascent TZs first assemble on basal bodies
during early pre-meiotic G2 phase (spermatocytes), anchor directly to the plasma membrane.
This is followed by establishment of cilium-membrane connections and a
ciliary membrane that persists during meiosis \citep{riparbelli2012assembly}
through a poorly understood mechanism
involving microtubule rearrangements \citep{gottardo2013cilium}.
Ciliogenesis begins with the assembly of TZ proteins at the tips of
the nascent basal body.
This is followed by the anchoring of the TZ to the membrane and
establishment of a compartmentalized space bounded by the ciliary membrane
and the TZ.
TZ maturation has been described in
\textit{Paramecium} \citep{aubusson2015transition}
\textit{Caenorrhabditis elegans} \citep{serwas2017centrioles} and
\textit{D. melanogaster} \citep{gottardo2013cilium}
and involves TZ elongation.
Possibly, this is similar to the initial steps of TZ assembly and reorganization
in well-studied vertebrate cilia.

Here, we show that the lipid phosphatidylinositol 4,5-bisphosphate (\PIP{})
is essential for proper TZ maturation in developing \textit{Drosophila melanogaster}
male germ cells.
Reduction of cellular \PIP{} by expression of the bacterial PIP phosphatase SigD
or by inactivation of the PIP 5-kinase Skittles (Sktl) that localizes to cilia
induces longer than normal TZs.
These hyperelongated TZs show hallmarks of aberrant function similar to those
found in previously studied TZ mutants, including loss of cilium-membrane associations.
We also find that the \textit{onion rings} allele of the
ciliopathy-associated \textit{EXOC8} gene encoding the human Exo84
decouples TZ hyperelongation from loss of cilium-membrane contacts.
Our results suggest that the plasma membrane lipid \PIP{} regulates TZ length,
maturation and function, acting to support ciliogenesis.

@@

Since exocyst regulates membrane trafficking and cilium formation,
we examined whether cilium-associated membrane dynamics were affected in \sigd{}
or \textit{onr} mutants similar to \textit{dila; cby} mutants
\citep{vieillard2016transition}.
We found that while TZs in \sigd{} and \textit{onr} cells anchored properly
to the plasma membrane, they were unable to either establish or maintain
membrane connections, and were rendered cytoplasmic (Figure 5) similar to
\textit{dila; cby} mutants.
Although we were not able to determine the localization of Exo84 due to the
lack of available antibodies,
we found that Sec6 and Exo70 localized to the plasma membrane as expected,
but were not enriched at the cilium.
Our results suggest that the exocyst may act downstream of \PIP{} to
enhance cilium-membrane associations, and that TZ hyperelongation and loss of
membrane-cilium association are genetically separable phenotypes.

@@

In this study, we show that the lipid \PIP{} is essential for proper
transition zone maturation in the Drosophila male germline.
Reduction of cellular \PIP{} by expression of the bacterial PIP
phosphatase SigD or inactivation of the type I PIP kinase Skittles
causes developing TZs to elongate excessively.

We had previously shown that reduction of \PIP{} by SigD expression
affected flagellar biogenesis and axoneme structure
\citep{wei2008depletion}.
We also showed aberrant ``comet-shaped'' localization of Unc that
failed to assemble TZ-associated puncta in some cases.
Here, we show that this occurs prior to meiosis,
is not an artefact due to cytokinesis defects
and do not represent an inability of the axoneme to grow.

The type I phosphatidylinositol 4-phosphate kinase (PIPKI) Skittles
suppresses TZ elongation in \sigd{} cells in a cilium-autonomous
manner.
We also found that cells mutant for \textit{sktl} contained hyperelongated
TZs similar to SigD-expressing cells.
Sktl also localizes to centrioles.

@@

Members of the exocyst complex, such as Sec10 and Sec6, have been shown
to be important for cilium formation in cultured cell lines and zebrafish
\citep{zuo2009exocyst, lobo2017exocyst, seixas2016arl13b}.
A mutation in \textit{EXOC8} has been linked
to Joubert syndrome \citep{dixon2012exome}.
The exocyst is known to bind to \PIP{}, and we previously described that
the \textit{onr} allele of Drosophila \textit{Exo84} phenocopied
male germline defects observed in \sigd{}.

\section{Methods}
\subsection{Transgenic flies}
Drosophila stocks were cultured on cornmeal molasses agar medium at 25$^{\circ}$C
and 50\% humidity.
Stocks expressing $\beta$\textit{2tub}::SigD and $\beta$\textit{2tub}::YFP-Sktl
were described previously \citep{wei2008depletion, fabian2010phosphatidylinositol}.
Ana1-tdTomato (chromosome 2) and Cep290-GFP (chromosome 3) were provided
by T. Avidor-Reiss \citep{basiri2014migrating}.
Unc-GFP was originally provided by M. Kernan \citep{baker2004mechanosensory}.
Stocks expressing GFP-tagged Chibby, B9d1, B9d2, Cc2d2a and Mks1 were generously
provided by B. Durand \citep{enjolras2012drosophila, vieillard2016transition}.
The \textit{onr} mutant was described previously \citep{giansanti2015exocyst}.
Stocks for generating \textit{sktl}\textsuperscript{2.3} clones were originally provided by
A. Guichet \citep{gervais2008pip5k}.
\textit{w}\textsuperscript{1118} was used as the wild-type control.

@@

\subsection{Fluorescence microscopy}
For live imaging, testes were dissected in phosphate bufferd saline (PBS).
If staining for chromatin, intact testes were incubated in PBS with
Hoechst (1:5000) for 5 minutes at this point.
Testes were transferred to a polylysine-coated glass slide in a drop of PBS,
ruptured using a syringe needle and
squashed under a glass coverslip using Kimwipes.
The edges of the coverslip were sealed with nail polish
and the specimen was visualized immediately using a fluorescence microscope.
